---
title: "LAB 1"
author: "Joshua Lee"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks
   
## Task 1

1) Obtain the working directory.

```{r}
getwd()
```
## Task 2
   
1) Read in the DDT file
   
```{r}
ddt <- read.csv("DDT.csv")
head(ddt)
```
   
## Task 3

1) The qualitative variables in ddt are RIVER and SPECIES.

2) The quantitative variables in ddt are MILE, LENGTH, WIDTH, and DDT.


```{r}
species <- ddt$SPECIES
print(species)
```

3) There are 3 different SPECIES in the ddt data set.

4) Subset the ddt data set so that we see only the observations with SPECIES,  LMBASS and WEIGHT bigger than 800gms

```{r}
with(ddt, ddt[SPECIES == "LMBASS" & WEIGHT > 800,])
```
5) Subset the ddt data set so that we see only those observations where the RIVER is SCM   and DDT bigger than 4.0


```{r}
with(ddt, ddt[RIVER == "SCM" & DDT > 4.0,])

```
CLICKER QUESTIONS

1) Find the mean length of fish in the ddt data frame.

```{r}
l = ddt$LENGTH
mean(l)
```
2) What is the standard deviation of the weight of fish in the ddt data frame?

```{r}
w = ddt$WEIGHT
sd(w)
```
3) Is this the plot of LENGTH vs WEIGHT from the ddt data frame?

   A: Yes

```{r}
plot(l, w, col = 10:16, main = "Length vs Weight")
```

4) If v=1:20 what is the last value of v/20?

   A: Since the last value of v is 20, then v/20 would equal 1

## Task 4

1). Make a table of the rivers.

```{r}
riv = ddt$RIVER
table(riv)
```

2) Make a color coded barplot of the rivers.

```{r}
tableR = table(riv)
barplot(tableR, col = 10:16, main = "Rivers")
```

3) Make a table of the rivers crossed with the species of fish.

```{r}
table(riv, species)
```
4) Make a barplot of the rivers crossed with the species of fish.

```{r}
tableRxS = table(riv, species)
barplot(tableRxS, beside = TRUE, col = 10:16, main = "Rivers vs Species")
```

## Task 5

1) Make a pie chart of Fish SPECIES
```{r}
pie(table(species), main = "Species")
```
2) Make a pie chart of RIVERS

```{r}
pie(table(riv), main = "Rivers")
```

## Task 6

1) Make a boxplot of the variable DDT (I cut off two points)

```{r}
boxplot(ddt$DDT, xlab = "DDT", ylim = c(0,400))
```

2) Make a boxplot of the variable WEIGHT

```{r}
boxplot(w, xlab = "WEIGHT")
```

3) Make a boxplot of the variable LENGTH

```{r}
boxplot(l, xlab = "LENGTH")
```

## Task 7

1) Make a coplot LENGTH vs WEIGHT given RIVER

```{r}
RIVERS = ddt$RIVER
coplot(l~w | RIVERS, col = 10:16, xlab = "Weight", ylab = "Length")

```

## Task 8

Creating the 3 plots with ggplot

1) Boxplot
```{r}
library(ggplot2)
box1 = ggplot(data = ddt, aes(x = SPECIES, y = WEIGHT))
box1 = box1 + geom_boxplot(aes(fill = RIVER)) + labs(title = "Joshua Lee")
box1
```

2) Violin Plot

```{r}
violin1 = ggplot(data = ddt, aes(x = RIVER, y = LENGTH))
violin1 = violin1 + geom_violin(aes(fill = SPECIES)) + labs(title = "Joshua Lee")
violin1
```

3) Scatter plot

```{r}
scatter1 = ggplot(data = ddt, aes(x = WEIGHT, y = LENGTH), aes(fill = SPECIES))
scatter1 = scatter1 + geom_point(aes(colour = SPECIES)) + labs(title = "Joshua Lee")
scatter1
```

