---
title: "LAB 2"
author: "Joshua Lee"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

Obtain the working directory
```{r}
getwd()
```
## Task 2

Reading in the data and printing the first 6 lines

```{r}
epagas = read.csv("EPAGAS.csv")
head(epagas)
```

## Task 3

Make the object mpg

```{r}
mpg = epagas$ï..MPG
head(mpg)
```

Transform mpg to z
```{r}
z = (mpg - mean(mpg)) / sd(mpg)
head(z)
print(paste0("mean = ", mean(z)), quote = F)
print(paste0("variance = ", var(z)), quote = F)
print(paste0("standard deviation = ", sd(z)), quote = F)
```

Possible outliers for mpg

```{r}
po = mpg[abs(z) >= 2 & abs(z) <= 3]
po
```

Outliers

```{r}
o = mpg[abs(z) >= 3]
o
```

Dotplot

```{r}
library(lattice)
mpgDataFCol = ifelse(abs(z) > 3, "red", 
                     ifelse(abs(z) >= 2 & abs(z) <= 3, "blue", "black"))
dotplot(mpg, col = mpgDataFCol, main = "Plot of mpg values", xlab = "Miles per Gallon")
```

## Task 4

1) Make a boxplot of the mpg variable
  a. Make the box black
  b. Put a notch where the median goes
  c. Put a title on the graph
  d. Make the plot horizontal
  
```{r}
boxplot(mpg, data = epagas, main = "Boxplot of MPG", col = "black", notch = T, horizontal = T)
```

2) Using Chebyshev's theorem predict the proportion of data within 2 standard deviation of the mean of the mpg data

Chebyshev's theorem states: at least 

$$ 1 - \frac{1}{k^2} $$

of the data will fall within k standard deviation of the mean.

When k = 2

```{r}
k = 2
cheb = 1 - 1 / (k^2)
cheb
```

3) Use R to calculate the exact proportion within 2 standard deviation of the mean

```{r}
sum(mpg > (mean(mpg)-(2 * sd(mpg))) & mpg < (mean(mpg)+(2 * sd(mpg)))) / length(mpg)
```
4) Does Chebyshev agree with the data?
  Ans: Nope
  
5) Now use the empirical rule, what proportion of the data should be within 2 standard deviations of the mean?
  Ans: The empirical rule states that 95% of the data should be within 2 standard deviations.

6) How well does it correspond?
  Ans: The value that we got is about 1% off from what the empirical rule states.
  
7) Is the Empirical rule valid in this case? Why?
  Ans: The empirical rule is valid in this case as the distribution follows closely with the       bell-shapedd curve of a normal distribution.
